Individualized Measures of Partisan Fairness and Utility in Redistricting (??)

### 2021-08-31 paper/paper.Rmd
start estimation section

From here onward we will restrict our attention to the case of two parties, so $\Parties=\{A,B\}$, since that will simplify the notation and exposition.
But the same concepts and definitions naturally carry over to the multiparty setting.

Intuitively, a districting plan is fair if it does not harm the voters of one party more than another, on average.
This is easily expressed using our notation above; the average fairness of $Q$ versus counterfactual $\tilde Q$ may be defined as $$
    F = \E[H\mid V(I) = A] - \E[H\mid V(I) = B];
$$obviously the ordering of $A$ and $B$ is immaterial.
Informally, this quantity is the difference in the probability that a voter of each party is represented by the other party but counterfactually would have been represented by their own party.

Generally, we will want to evaluate the fairness of a specific districting plan $q$ versus a distribution of counterfactuals.
These counterfactuals could arise from a posited probability distribution capturing the factors relevant to a non-partisan redistricting process, or they could be a set of alternative redistricting plans that were considered as part of the legislative process leading to a final enacted plan.
We adopt the notation $F(q)$ for this scenario: $$
    F(q) \dfeq \E[H\mid V(I) = A, Q=q] - \E[H\mid V(I) = B, Q=q],
$$ so that we recover $F=\E[F(Q)]$.

@@

# Estimating the Harm and Fairness Measures {#sec:estimating}

>>>>

From here onward we will restrict our attention to the case of two parties---concretely, $\Parties=\{\dem,\rep\}$---since that will simplify the notation and exposition.
But the same concepts and definitions naturally carry over to the multiparty setting.

Intuitively, a districting plan is fair if it does not harm the voters of one party more than another, on average.
This is easily expressed using our notation above; the average fairness of $Q$ versus counterfactual $\tilde Q$ may be defined as $$
    F = \E[H\mid V(I) = \dem] - \E[H\mid V(I) = \rep];
$$obviously the ordering of $\dem$ and $\rep$ is immaterial.
Informally, this quantity is the difference in the probability that a voter of each party is represented by the other party but counterfactually would have been represented by their own party.

Generally, we will want to evaluate the fairness of a specific districting plan $q$ versus a distribution of counterfactuals.
These counterfactuals could arise from a posited probability distribution capturing the factors relevant to a non-partisan redistricting process, or they could be a set of alternative redistricting plans that were considered as part of the legislative process leading to a final enacted plan.
We adopt the notation $F(q)$ for this scenario: $$
    F(q) \dfeq \E[H\mid V(I) = \dem, Q=q] - \E[H\mid V(I) = \rep, Q=q],
$$ so that we recover $F=\E[F(Q)]$.

@@

# Calculating the Harm and Fairness Measures {#sec:estimating}

The measures $H$ and $F(q)$ defined above require specification of a probability measure on elections and districting plans, $\pi$.
Depending on one's viewpoint, this is either a theoretical fiction which must be specified by a model, or an actual data-generating process we must aim to estimate.
Either way, a collection of past election data and an enacted districting plan alone are not enough to calculate $H$ and $F(q)$.

While not strictly necessary, modelling of $V$ and $\tilde Q$ independently will simplify the situation considerably.
Practically, this translates to the assumption that individual voters would vote for the same political party no matter the districting plan.
To the extent that voters reward or punish an incumbent party for the districting plan they adopt, or that an adopted districting plan affects which candidates run in a district, this assumption is violated.
But as our suggested model for election outcomes will not account for candidate factors or policy preferences, the marginal effect on our conclusions will be minimal.

## Modelling Districting Plans

## Modelling Election Outcomes
